### 2025-05-25 Slides/Presentation/slides.tex

% ECE4900W Communicating Engineering Solutions
% Author: Arturo Salinas-Aguayo <artjsalina5@uconn.edu>
\documentclass{beamer}
\usetheme{Warsaw}
\usecolortheme{dolphin}
\usefonttheme{professionalfonts}
\setbeamertemplate{navigation symbols}{}

\title[Engineering Communication]{Communicating Engineering Solutions in a Societal Context}
\subtitle{ECE 4900W – Summer 2025}
\author[Arturo Salinas]{Arturo Salinas-Aguayo}
\institute[UConn]{University of Connecticut\\College of Engineering}
\date{\today}

\AtBeginSection[]{
  \begin{frame}{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\begin{frame}
	\titlepage
\end{frame}

\begin{frame}{Outline}
	\tableofcontents[pausesections]
\end{frame}

\section{Introduction}

\begin{frame}{Why This Course Matters}
	\begin{itemize}
		\item Engineers must communicate their solutions clearly.
		\item Societal impact matters: ethics, sustainability, inclusion.
		\item Writing and presenting ideas is just as important as building them.
	\end{itemize}
\end{frame}

\section{Course Components}

\begin{frame}{Major Assignments}
	\begin{itemize}
		\item \textbf{Paper One}: Technology Comparison (Draft + Final)
		\item \textbf{Paper Two}: Societal Impacts (Draft + Final)
		\item \textbf{Oral Presentation} based on Paper Two
		\item \textbf{Group Discussions} and \textbf{Exercises}
	\end{itemize}
\end{frame}

\subsection{Grading Breakdown}
\begin{frame}{Assessment Breakdown}
	\begin{itemize}
		\item Drafts \& Finals: \alert{63\%} combined
		\item Oral Presentation: \alert{14\%}
		\item Exercises \& Discussions: \alert{18\%}
		\item Peer Feedback: \alert{5\%}
	\end{itemize}
	\vfill
	\begin{block}{Rubrics Available}
		Check HuskyCT for detailed rubrics and formatting guides.
	\end{block}
\end{frame}

\section{Tools and Strategies}

\begin{frame}{Tips for Success}
	\begin{itemize}
		\item Start drafts early — use feedback!
		\item Visit the Writing Center
		\item Use IEEE citation style
		\item Write for clarity, audience, and purpose
	\end{itemize}
\end{frame}

\section{Summary}

\begin{frame}{Summary}
	\begin{itemize}
		\item \alert{Writing is Engineering}
		\item \alert{You will write, revise, and present}
		\item This course prepares you for communication in your future work
	\end{itemize}
	\vfill
	\begin{block}{Outlook}
		Capstone-ready communication skills that make you stand out.
	\end{block}
\end{frame}

\end{document}